---
title: "MDS Emotion and Mood"
author: "Monica Kullar"
date: "2/23/2024"
output: html_document
---
##################################################
##                  DEMOGRAPHICS                ##
##################################################

#assess risk for groupings
```{r}
library(reshape)
library(tidyverse)
library(dplyr)

myriad <- read.csv(file = "~/Desktop/Projects/Data/Myriad1B/Myriad1b_AgeGender.csv", sep=',', na.strings=c(""), header = T, stringsAsFactors = FALSE, colClasses = "character")
sapply(myriad, class)
myriad <- as.data.frame(myriad)
myriad$PtcptID <- as.factor(myriad$PtcptID)
myriad$Age <- as.numeric(myriad$Age)
myriad$Gender <- as.numeric(myriad$Gender)

#qualtrics data
myriadAgeDOBSex <- read.csv(file = "~/Desktop/Projects/Data/Myriad1B/myriadDOB_Sex.csv", sep=',', na.strings=c(""), header = T, stringsAsFactors = FALSE, colClasses = "character")
shortdemvars <- c("PtcptID", "Age", "Gender")
myriad <- myriad[shortdemvars]
myriad <- na.omit(myriad)
rownames(myriad) <- 1:nrow(myriad)
myriad$PtcptID <- as.numeric(myriad$PtcptID)
#remove participants who did not complete SPAM tasks
myriad<-myriad[!(myriad$PtcptID==54),]
myriad<-myriad[!(myriad$PtcptID==60),]
myriad<-myriad[!(myriad$PtcptID==76),]
rownames(myriad) <- 1:nrow(myriad)

## CHECKING CES-D, DEPRESSION RISK
cesD <- read.csv(file = "~/Desktop/Projects/Data/Myriad1B/CESD.csv", sep=',', na.strings=c(""), header = T, stringsAsFactors = FALSE, colClasses = "character")
cesD <- cbind(cesD[,1], cesD[ , grepl( "CES" , names(cesD)) ])
cesD <- cesD %>% 
  rename_with(.cols = 1, ~"PtcptID")
cesD$PtcptID <- as.numeric(cesD$PtcptID)
#missing IDs 46 & 53 & 75 from this datasheet...all else present. Merge in to add 46/53 into the CES-D data using their myriadQualtrics CES-D data
cesD465375 <- read.csv(file = "~/Desktop/Projects/Data/Myriad1B/myriadQualtrics_rawCESD.csv", sep=',', na.strings=c(""), header = T, stringsAsFactors = FALSE, colClasses = "character")
cesD465375 <- cesD465375[-(1:2),]
rownames(cesD465375) <- 1:nrow(cesD465375)
cesD465375 <- cesD465375 %>% filter(PtcptID == "46" |PtcptID ==  "53"|PtcptID ==  "75")
cesD465375 <- cbind(cesD465375[,20], cesD465375[ , grepl( "CES" , names(cesD465375)) ])
cesD465375 <- cesD465375 %>% 
  rename_with(.cols = 1, ~"PtcptID")
cesD465375$PtcptID <- as.numeric(cesD465375$PtcptID)
#make CES-D numeric for scoring
cesD465375$CESD_1 <- as.numeric(cesD465375$CESD_1)
cesD465375$CESD_2 <- as.numeric(cesD465375$CESD_2)
cesD465375$CESD_3 <- as.numeric(cesD465375$CESD_3)
cesD465375$CESD_4 <- as.numeric(cesD465375$CESD_4)
cesD465375$CESD_5 <- as.numeric(cesD465375$CESD_5)
cesD465375$CESD_6 <- as.numeric(cesD465375$CESD_6)
cesD465375$CESD_7 <- as.numeric(cesD465375$CESD_7)
cesD465375$CESD_8 <- as.numeric(cesD465375$CESD_8)
cesD465375$CESD_9 <- as.numeric(cesD465375$CESD_9)
cesD465375$CESD_10 <- as.numeric(cesD465375$CESD_10)
cesD465375$CESD_11 <- as.numeric(cesD465375$CESD_11)
cesD465375$CESD_12 <- as.numeric(cesD465375$CESD_12)
cesD465375$CESD_13 <- as.numeric(cesD465375$CESD_13)
cesD465375$CESD_14 <- as.numeric(cesD465375$CESD_14)
cesD465375$CESD_15 <- as.numeric(cesD465375$CESD_15)
cesD465375$CESD_16 <- as.numeric(cesD465375$CESD_16)
cesD465375$CESD_17 <- as.numeric(cesD465375$CESD_17)
cesD465375$CESD_18 <- as.numeric(cesD465375$CESD_18)
cesD465375$CESD_19 <- as.numeric(cesD465375$CESD_19)
cesD465375$CESD_20 <- as.numeric(cesD465375$CESD_20)
cesD465375 <- na.omit(cesD465375)
#calculate total given raw non-reversed scores at 4,8,12,16
#row 1 (53 t1) = 19
#row 2 (46) = 11
cesD465375$CESD_TOTAL <- c(2, 19, 11)
#merge full available cesD dataset together
cesD <- rbind(cesD, cesD465375)
rownames(cesD) <- 1:nrow(cesD)
#merge demographics info to cesD info for participants present for MoodSPAM
cesDmyriad <- merge(myriad, cesD, by = "PtcptID")
rownames(cesDmyriad) <- 1:nrow(cesDmyriad)
#make cesD final scores all numeric, and participantID factor
cesDmyriad$PtcptID <- as.factor(cesDmyriad$PtcptID)
cesDmyriad$CESD_1 <- as.numeric(cesDmyriad$CESD_1)
cesDmyriad$CESD_2 <- as.numeric(cesDmyriad$CESD_2)
cesDmyriad$CESD_3 <- as.numeric(cesDmyriad$CESD_3)
cesDmyriad$CESD_4 <- as.numeric(cesDmyriad$CESD_4)
cesDmyriad$CESD_5 <- as.numeric(cesDmyriad$CESD_5)
cesDmyriad$CESD_6 <- as.numeric(cesDmyriad$CESD_6)
cesDmyriad$CESD_7 <- as.numeric(cesDmyriad$CESD_7)
cesDmyriad$CESD_8 <- as.numeric(cesDmyriad$CESD_8)
cesDmyriad$CESD_9 <- as.numeric(cesDmyriad$CESD_9)
cesDmyriad$CESD_10 <- as.numeric(cesDmyriad$CESD_10)
cesDmyriad$CESD_11 <- as.numeric(cesDmyriad$CESD_11)
cesDmyriad$CESD_12 <- as.numeric(cesDmyriad$CESD_12)
cesDmyriad$CESD_13 <- as.numeric(cesDmyriad$CESD_13)
cesDmyriad$CESD_14 <- as.numeric(cesDmyriad$CESD_14)
cesDmyriad$CESD_15 <- as.numeric(cesDmyriad$CESD_15)
cesDmyriad$CESD_16 <- as.numeric(cesDmyriad$CESD_16)
cesDmyriad$CESD_17 <- as.numeric(cesDmyriad$CESD_17)
cesDmyriad$CESD_18 <- as.numeric(cesDmyriad$CESD_18)
cesDmyriad$CESD_19 <- as.numeric(cesDmyriad$CESD_19)
cesDmyriad$CESD_20 <- as.numeric(cesDmyriad$CESD_20)

test0 <- melt(cesDmyriad[c("PtcptID", "CESD_TOTAL")], id= "PtcptID")
test0$value <- as.numeric(test0$value)
test0$PtcptID <- as.factor(test0$PtcptID)
ggplot(test0, aes(x=variable, y=value)) + geom_point() + theme_classic() + geom_hline(yintercept=16, col="red") + ggtitle("CES-D from Spreadsheet")

ggplot(test0, aes(x=value)) + geom_histogram(aes(fill = value < 16)) + theme_classic() +
  scale_fill_manual(values = c("coral", "light green")) + ggtitle("Depression Risk via CES-D")

#CHECK DEMOGRAPHICS OF CES-D GROUPS
# N=69
myriad$PtcptID <- as.factor(myriad$PtcptID)
test0$PtcptID <- as.factor(test0$PtcptID)
test0dem <- merge(test0, myriad, by = "PtcptID")

#Add a group column for low/at risk
test0dem$Group <- with(test0dem, ifelse(value<16, "Low Risk", "At Risk"))
sapply(test0dem, class)
test0dem$Group <- as.factor(test0dem$Group)
```

##################################################
##                ALL EMOTION TERMS             ##
##################################################

#data setup
```{r}
library(tidyverse)

distance <- read.table("~/Data/spam_emotion_merge.txt", sep = "\t", header = TRUE) #must delete first line of C: saving location in txt file
short <- c("Subject", "Comp1", "Comp2", "Distance")
distance <- distance[short]
distance <- na.omit(distance)
distance$Subject <- as.factor(distance$Subject)
distance$Comp1 <- as.factor(distance$Comp1)
distance$Comp2 <- as.factor(distance$Comp2)

levels(distance$Comp2)
posSPAMmood <- c("CALM", "CAREFREE", "CHEERFUL", "COMPOSED", "CONTENT", "EXCITED", "HAPPY", "JOYFUL", "LIVELY", "MANIC", "OPTIMISTIC", "PASSIONATE", "RELAXED", "SATISFIED", "SERENE")
negSPAMmood <- c("ANXIOUS", "BITTER", "BORED",  "DEPRESSED", "DESPERATE", "FED UP", "GLOOMY", "GUILTY", "IRRITABLE", "MISERABLE", "PANICKY", "PESSIMISTIC", "SAD", "STRESSED", "TENSE", "UNEASY")

distance <- distance %>% 
  rename_with(.cols = 1, ~"PtcptID")

distancetest <- merge(distance, test0dem[,c(1,6)], by= "PtcptID")
distancetest$termpairs <- paste(distancetest$Comp1, distancetest$Comp2, sep= "-")
```

# Loops for individual matrices, individual configuration plots
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)

rownames(distancetest) <- 1:nrow(distancetest)

# INDIVIDUAL MATRIX CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #save all files in Imp folder for use at individual level
    savefile <- paste0 ("~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/", i, ".csv", sep = "")
    write.csv(SpamMatrixtemp, file = savefile, row.names = TRUE)
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}

# INDIVIDUAL GGSCATTER CONFIGURATION PLOT CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #create their mds plot
    spam_mdstemp <-  mds(delta = SpamMatrixtemp, ndim = 2, type = "ordinal")
  #simple clean readable config plot
  plottemp <- ggplot() + geom_point(data = as.data.frame(spam_mdstemp$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdstemp$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(SpamMatrixtemp)) + 
  labs(title = "MDS representation of Emotion") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()

    #save all files in Imp folder for use at individual level
    savefile <- paste0 (i, ".jpg", sep = "")
    ggsave(file=savefile, plot=plottemp, path = "~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/indivconfigplotsEMO/")
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}
```

# Average Low and At Risk Group analysis of matrices (averaging matrices, stress, plots, variance diff)
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(stats)

#Average matrices for low and at risk groups by SUBFOLDER OF PRESORTED GROUPS.
    ## LOW RISK ##
#pull all the csv files from the subfolder
filenamesLR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/lowrisk", pattern="*.csv", full.names=TRUE)
#read them in as csv, then as matrices which is what i want
ldfLR <- lapply(filenamesLR, read.csv)

resLR <- lapply(ldfLR, as.matrix)

#make each one have their mood terms as row names via lapply.
resLR <- lapply(ldfLR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
#matrix average LOW RISK
AvgMatrixLoRisk <- Reduce(`+`, resLR)/length(resLR)

    ## AT RISK ##
filenamesHR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/atrisk", pattern="*.csv", full.names=TRUE)
ldfHR <- lapply(filenamesHR, read.csv)
resHR <- lapply(ldfHR, as.matrix)
#make each one have their mood terms as row names via lapply.
resHR <- lapply(ldfHR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
#matrix average AT RISK
AvgMatrixHiRisk <- Reduce(`+`, resHR)/length(resHR)

#Assess group matrices 
#analyze configuration plots, stress values, normalized stress, and k-means clustering of outputs.

#Low Risk
spam_mdsLR <-  mds(delta = AvgMatrixLoRisk, ndim = 2, type = "ordinal" )
spam_mdsLR
spam_mdsLR$stress #the stress value = .04 GOOD
dhat_matrix <-  as.matrix(spam_mdsLR$dhat)
d_matrix <-  as.matrix(spam_mdsLR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressLR  <-  nominator/denominator
normalized_stressLR #the normalized stress value = .002

#At Risk
spam_mdsHR <-  mds(delta = AvgMatrixHiRisk, ndim = 2, type = "ordinal" )
spam_mdsHR
spam_mdsHR$stress #the stress value = .05 GOOD
dhat_matrix <-  as.matrix(spam_mdsHR$dhat)
d_matrix <-  as.matrix(spam_mdsHR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressHR  <-  nominator/denominator
normalized_stressHR #the normalized stress value = .003

# STATISTIC FOR ASSESSING BORED OUTLIER
checkingoutliers <- as.data.frame(spam_mdsLR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)

checkingoutliers <- as.data.frame(spam_mdsHR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)

#Group plots by risk status
plot(spam_mdsLR)
plot(spam_mdsHR)
#Low risk
LRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsLR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsLR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixLoRisk)) + 
  labs(title = "MDS representation of Emotion: Low Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()
    ggsave(file="LRplotEMO.jpg", width=8, height=8, plot=LRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
#At risk
HRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsHR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsHR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixHiRisk)) + 
  labs(title = "MDS representation of Emotion: At Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic() +
  scale_y_continuous(limits = c(-.5, .4)) #make same as low risk plot Y (X is already same)
    ggsave(file="HRplotEMO.jpg",  width=8, height=8, plot=HRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
    
#compare spread of both dimensions.
LRscores <- as.data.frame(spam_mdsLR$conf)
mean(LRscores$D1)
sd(LRscores$D1)
mean(LRscores$D2)
sd(LRscores$D2)

HRscores <- as.data.frame(spam_mdsHR$conf)
mean(HRscores$D1)
sd(HRscores$D1)
mean(HRscores$D2)
sd(HRscores$D2)

#t test for mean differences
t.test(LRscores$D1, HRscores$D1)
#t = -2.6686e-16, df = 59.955, p-value = 1
t.test(LRscores$D2, HRscores$D2)
#t = -4.859e-15, df = 49.249, p-value = 1

#-test of two variances
var.test(LRscores$D1, HRscores$D1)
var.test(LRscores$D2, HRscores$D2)

LRscores$Group <- "LowRisk"
HRscores$Group <- "AtRisk"

LRHR <- rbind(LRscores, HRscores)
#plot the spread
#boxplot(D2 ~ Group, data = LRHR)
ggplot(LRHR, aes(x=Group, y=D1, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 1 between Groups") +
  theme_classic()
ggplot(LRHR, aes(x=Group, y=D2, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 2 between Groups") +
  theme_classic()

#within-groups variance of dimensions
var.test(LRscores$D1, LRscores$D2)
var.test(HRscores$D1, HRscores$D2)
```

#Clustering, Cuts
```{r}
library(stats)
library(dplyr)
library(ggplot2)
library(ggpubr)
mdsLR <- spam_mdsLR$conf
mdsHR <- spam_mdsHR$conf

# Determine number of clusters
wss <- (nrow(mdsLR)-1)*sum(apply(mdsLR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsLR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="Low Risk Group") #strongly 2

wss <- (nrow(mdsHR)-1)*sum(apply(mdsHR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsHR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="At Risk Group") #also strongly 2

# Ward Hierarchical Clustering
d <- dist(mdsLR, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward.D")
plot(fit) # display dendogram
groups <- cutree(fit, k=4) # cut tree into N clusters
# draw dendogram with red borders around the N clusters
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")

d <- dist(mdsHR, method = "euclidean")
fit <- hclust(d, method="ward.D")
plot(fit)
groups <- cutree(fit, k=4)
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")
```

##################################################
##    EMOTION TERMS, WITHOUT 'BORED' OUTLIER    ##
##################################################

#data setup
```{r}
distance <- read.table("~/ownCloud/PhD/Projects/Data/Myriad1B/spam_emotion/spam_emotion_merge.txt", sep = "\t", header = TRUE) #must delete first line of C: saving location in txt file
short <- c("Subject", "Comp1", "Comp2", "Distance")
distance <- distance[short]
distance <- na.omit(distance)
sapply(distance, class)
distance$Subject <- as.factor(distance$Subject)
distance$Comp1 <- as.factor(distance$Comp1)
distance$Comp2 <- as.factor(distance$Comp2)

posSPAMmood <- c("CALM", "CAREFREE", "CHEERFUL", "COMPOSED", "CONTENT", "EXCITED", "HAPPY", "JOYFUL", "LIVELY", "MANIC", "OPTIMISTIC", "PASSIONATE", "RELAXED", "SATISFIED", "SERENE")
negSPAMmood <- c("ANXIOUS", "BITTER", "BORED",  "DEPRESSED", "DESPERATE", "FED UP", "GLOOMY", "GUILTY", "IRRITABLE", "MISERABLE", "PANICKY", "PESSIMISTIC", "SAD", "STRESSED", "TENSE", "UNEASY")
distance <- distance %>% 
  rename_with(.cols = 1, ~"PtcptID")
distancetest <- merge(distance, test0dem[,c(1,6)], by= "PtcptID")
distancetest$termpairs <- paste(distancetest$Comp1, distancetest$Comp2, sep= "-")

    # REMOVE BORED (outlier term)
distancetest <- distancetest[!grepl("BORED", distancetest$Comp1),]
distancetest <- distancetest[!grepl("BORED", distancetest$Comp2),]
```

# Loops for individual matrices, individual configuration plots
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)

rownames(distancetest) <- 1:nrow(distancetest)

# INDIVIDUAL MATRIX CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #save all files in Imp folder for use at individual level
    savefile <- paste0 ("~/ownCloud/PhD/Projects/Data/Myriad1B/indivmatricesEMO/", i, ".csv", sep = "")
    write.csv(SpamMatrixtemp, file = savefile, row.names = TRUE)
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}

# INDIVIDUAL GGSCATTER CONFIGURATION PLOT CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #create their mds plot
    spam_mdstemp <-  mds(delta = SpamMatrixtemp, ndim = 2, type = "ordinal")
  #simple clean readable config plot
  plottemp <- ggplot() + geom_point(data = as.data.frame(spam_mdstemp$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdstemp$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(SpamMatrixtemp)) + 
  labs(title = "MDS representation of Emotion") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()

    #save all files in Imp folder for use at individual level
    savefile <- paste0 (i, ".jpg", sep = "")
    ggsave(file=savefile, plot=plottemp, path = "~/ownCloud/PhD/Projects/Data/Myriad1B/indivmatricesEMO/indivconfigplotsEMO_noBored")
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}
```

# Average Low and At Risk Group analysis of matrices (averaging matrices, stress, plots, variance diff)
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(stats)

#Average matrices for low and at risk groups by SUBFOLDER OF PRESORTED GROUPS.
    ## LOW RISK ##
#pull all the csv files from the subfolder
filenamesLR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/lowrisk_noBored", pattern="*.csv", full.names=TRUE)
ldfLR <- lapply(filenamesLR, read.csv)

resLR <- lapply(ldfLR, as.matrix)

#make each one have their mood terms as row names via lapply.
resLR <- lapply(ldfLR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
#matrix average LOW RISK
AvgMatrixLoRisk <- Reduce(`+`, resLR)/length(resLR)

    ## AT RISK ##
filenamesHR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatricesEMO/atrisk_noBored", pattern="*.csv", full.names=TRUE)
ldfHR <- lapply(filenamesHR, read.csv)
resHR <- lapply(ldfHR, as.matrix)
#make each one have their mood terms as row names via lapply.
resHR <- lapply(ldfHR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
AvgMatrixHiRisk <- Reduce(`+`, resHR)/length(resHR)

#Assess group matrices

#Low Risk
spam_mdsLR <-  mds(delta = AvgMatrixLoRisk, ndim = 2, type = "ratio" )
spam_mdsLR
spam_mdsLR$stress #the stress value = .12
dhat_matrix <-  as.matrix(spam_mdsLR$dhat)
d_matrix <-  as.matrix(spam_mdsLR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressLR  <-  nominator/denominator
normalized_stressLR #the normalized stress value = .02

#At Risk
spam_mdsHR <-  mds(delta = AvgMatrixHiRisk, ndim = 2, type = "ratio" )
spam_mdsHR
spam_mdsHR$stress #the stress value = .14
dhat_matrix <-  as.matrix(spam_mdsHR$dhat)
d_matrix <-  as.matrix(spam_mdsHR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressHR  <-  nominator/denominator
normalized_stressHR #the normalized stress value = .02

#compare spread of both dimensions.
LRscores <- as.data.frame(spam_mdsLR$conf)
mean(LRscores$D1)
sd(LRscores$D1)
mean(LRscores$D2)
sd(LRscores$D2)

EmoLRscores <- LRscores

HRscores <- as.data.frame(spam_mdsHR$conf)
mean(HRscores$D1)
sd(HRscores$D1)
mean(HRscores$D2)
sd(HRscores$D2)

EmoHRscores <- HRscores

#Group plots by risk status
plot(spam_mdsLR)
plot(spam_mdsHR)
#Low risk
LRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsLR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsLR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixLoRisk)) + 
  labs(title = "MDS representation of Emotion: Low Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()
    ggsave(file="LRplotEMOnoBoredRatio.jpg", width=8, height=8, plot=LRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
#At risk
HRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsHR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsHR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixHiRisk)) + 
  labs(title = "MDS representation of Emotion: At Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic() +
  scale_y_continuous(limits = c(-.5, .4)) #make same as low risk plot Y (X is already same)
    ggsave(file="HRplotEMOnoBoredRatio.jpg", width=8, height=8, plot=HRplot, path = "~/Desktop/Projects/Data/Myriad1B/")

#t test for mean differences
t.test(LRscores$D1, HRscores$D1)
#t = -2.6686e-16, df = 59.955, p-value = 1
t.test(LRscores$D2, HRscores$D2)
#t = -4.859e-15, df = 49.249, p-value = 1

#ð¹-test of two variances.
var.test(LRscores$D1, HRscores$D1)
var.test(LRscores$D2, HRscores$D2)

LRscores$Group <- "LowRisk"
HRscores$Group <- "AtRisk"

LRHR <- rbind(LRscores, HRscores)
#plot the spread
#boxplot(D2 ~ Group, data = LRHR)
ggplot(LRHR, aes(x=Group, y=D1, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 1 between Groups") +
  theme_classic()
ggplot(LRHR, aes(x=Group, y=D2, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 2 between Groups") +
  theme_classic()

#within-groups variance of dimensions
var.test(LRscores$D1, LRscores$D2)
var.test(HRscores$D1, HRscores$D2)
```

#Clustering, Cuts
```{r}
library(stats)
library(dplyr)
library(ggplot2)
library(ggpubr)
mdsLR <- spam_mdsLR$conf
mdsHR <- spam_mdsHR$conf

# Determine number of clusters
wss <- (nrow(mdsLR)-1)*sum(apply(mdsLR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsLR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="Low Risk Group") #strongly 2

wss <- (nrow(mdsHR)-1)*sum(apply(mdsHR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsHR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="At Risk Group") #also strongly 2

# Ward Hierarchical Clustering
d <- dist(mdsLR, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward.D")
plot(fit) # display dendogram
groups <- cutree(fit, k=4) # cut tree into N clusters
# draw dendogram with red borders around the N clusters
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")

d <- dist(mdsHR, method = "euclidean")
fit <- hclust(d, method="ward.D")
plot(fit)
groups <- cutree(fit, k=4)
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")
```

##################################################
##                ALL MOOD TERMS                ##
##################################################

#data setup
```{r}
library(tidyverse)

distance <- read.table("~/Data/spammood75merge.txt", sep = "\t", header = TRUE)
short <- c("Subject", "Comp1", "Comp2", "Distance")
distance <- distance[short]
distance <- na.omit(distance)
sapply(distance, class)
distance$Subject <- as.factor(distance$Subject)
distance$Comp1 <- as.factor(distance$Comp1)
distance$Comp2 <- as.factor(distance$Comp2)

levels(distance$Comp2)
posSPAMmood <- c("CALM", "CAREFREE", "CHEERFUL", "COMPOSED", "CONTENT", "EXCITED", "HAPPY", "JOYFUL", "LIVELY", "MANIC", "OPTIMISTIC", "PASSIONATE", "RELAXED", "SATISFIED", "SERENE")
negSPAMmood <- c("ANXIOUS", "BITTER", "BORED",  "DEPRESSED", "DESPERATE", "FED UP", "GLOOMY", "GUILTY", "IRRITABLE", "MISERABLE", "PANICKY", "PESSIMISTIC", "SAD", "STRESSED", "TENSE", "UNEASY")

distance <- distance %>% 
  rename_with(.cols = 1, ~"PtcptID")
distancetest$termpairs <- paste(distancetest$Comp1, distancetest$Comp2, sep= "-")
```

# Loops for individual matrices, individual configuration plot
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)

rownames(distancetest) <- 1:nrow(distancetest)

# INDIVIDUAL MATRIX CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #save all files in Imp folder for use at individual level
    savefile <- paste0 ("~/Desktop/Projects/Data/Myriad1B/indivmatrices/", i, ".csv", sep = "")
    write.csv(SpamMatrixtemp, file = savefile, row.names = TRUE)
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}

# INDIVIDUAL GGSCATTER CONFIGURATION PLOT CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #create their mds plot
    spam_mdstemp <-  mds(delta = SpamMatrixtemp, ndim = 2, type = "ordinal")
  #simple clean readable config plot
  plottemp <- ggplot() + geom_point(data = as.data.frame(spam_mdstemp$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdstemp$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(SpamMatrixtemp)) + 
  labs(title = "MDS representation of Mood") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()

    #save all files in Imp folder for use at individual level
    savefile <- paste0 (i, ".jpg", sep = "")
    ggsave(file=savefile, plot=plottemp, path = "~/Desktop/Projects/Data/Myriad1B/indivmatrices/indivconfigplots/")
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}
```

# Average Low and At Risk Group analysis of matrices (averaging matrices, stress, plots, variance diff)
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(stats)

#Average matrices for low and at risk groups by SUBFOLDER OF PRESORTED GROUPS.
    ## LOW RISK ##
#pull all the csv files from the subfolder
filenamesLR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatrices/lowrisk", pattern="*.csv", full.names=TRUE)
ldfLR <- lapply(filenamesLR, read.csv)
resLR <- lapply(ldfLR, as.matrix)
#make each one have their mood terms as row names via lapply.
resLR <- lapply(ldfLR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
#matrix average LOW RISK
AvgMatrixLoRisk <- Reduce(`+`, resLR)/length(resLR)

    ## AT RISK ##
filenamesHR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatrices/athrisk", pattern="*.csv", full.names=TRUE)
ldfHR <- lapply(filenamesHR, read.csv)
resHR <- lapply(ldfHR, as.matrix)
#make each one have their mood terms as row names via lapply.
resHR <- lapply(ldfHR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
AvgMatrixHiRisk <- Reduce(`+`, resHR)/length(resHR)

#Assess group matrices 
#analyze configuration plots, stress values, normalized stress, and k-means clustering of outputs.

#Low Risk
spam_mdsLR <-  mds(delta = AvgMatrixLoRisk, ndim = 2, type = "ordinal" )
spam_mdsLR
spam_mdsLR$stress #the stress value = .07 GOOD
dhat_matrix <-  as.matrix(spam_mdsLR$dhat)
d_matrix <-  as.matrix(spam_mdsLR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressLR  <-  nominator/denominator
normalized_stressLR #the normalized stress value = .005

#At Risk
spam_mdsHR <-  mds(delta = AvgMatrixHiRisk, ndim = 2, type = "ordinal" )
spam_mdsHR
spam_mdsHR$stress #the stress value = .05 GOOD
dhat_matrix <-  as.matrix(spam_mdsHR$dhat)
d_matrix <-  as.matrix(spam_mdsHR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressHR  <-  nominator/denominator
normalized_stressHR #the normalized stress value = .002

# STATISTIC FOR CONFIRMING BORED OUTLIER
checkingoutliers <- as.data.frame(spam_mdsLR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)
z_scores

checkingoutliers <- as.data.frame(spam_mdsHR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)
z_scores


#Group plots by risk status
plot(spam_mdsLR)
plot(spam_mdsHR)
#Low risk
LRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsLR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsLR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixLoRisk)) + 
  labs(title = "MDS representation of Mood: Low Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()
    ggsave(file="LRplot.jpg", plot=LRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
#At risk
HRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsHR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsHR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixHiRisk)) + 
  labs(title = "MDS representation of Mood: At Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic() +
  scale_y_continuous(limits = c(-.5, .4)) #make same as low risk plot Y (X is already same)
    ggsave(file="HRplot.jpg", plot=HRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
    
#compare spread of both dimensions.
LRscores <- as.data.frame(spam_mdsLR$conf)
mean(LRscores$D1)
sd(LRscores$D1)
mean(LRscores$D2)
sd(LRscores$D2)

HRscores <- as.data.frame(spam_mdsHR$conf)
mean(HRscores$D1)
sd(HRscores$D1)
mean(HRscores$D2)
sd(HRscores$D2)

#t test for mean differences
t.test(LRscores$D1, HRscores$D1)
#t = -2.6686e-16, df = 59.955, p-value = 1
t.test(LRscores$D2, HRscores$D2)
#t = -4.859e-15, df = 49.249, p-value = 1

#ð¹-test of two variances
var.test(LRscores$D1, HRscores$D1)
var.test(LRscores$D2, HRscores$D2)

LRscores$Group <- "LowRisk"
HRscores$Group <- "AtRisk"

LRHR <- rbind(LRscores, HRscores)
#plot the spread
#boxplot(D2 ~ Group, data = LRHR)
ggplot(LRHR, aes(x=Group, y=D1, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 1 between Groups") +
  theme_classic()
ggplot(LRHR, aes(x=Group, y=D2, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 2 between Groups") +
  theme_classic()

#within-groups variance of dimensions
var.test(LRscores$D1, LRscores$D2)
var.test(HRscores$D2, HRscores$D2)

```

#Clustering, Cuts
```{r}
library(stats)
library(dplyr)
library(ggplot2)
library(ggpubr)
mdsLR <- spam_mdsLR$conf
mdsHR <- spam_mdsHR$conf

# Determine number of clusters
wss <- (nrow(mdsLR)-1)*sum(apply(mdsLR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsLR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="Low Risk Group") #strongly 2

wss <- (nrow(mdsHR)-1)*sum(apply(mdsHR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsHR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="At Risk Group") #also strongly 2

# Ward Hierarchical Clustering
d <- dist(mdsLR, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=4) # cut tree into N clusters
# draw dendogram with red borders around the N clusters
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")

d <- dist(mdsHR, method = "euclidean")
fit <- hclust(d, method="ward")
plot(fit)
groups <- cutree(fit, k=4)
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")
```

##################################################
##       MOOD TERMS, WITHOUT 'BORED' OUTLIER    ##
##################################################

#data setup
```{r}
library(tidyverse)

distance <- read.table("~/ownCloud/PhD/Projects/Data/Myriad1B/spam_mood/spammood75merge.txt", sep = "\t", header = TRUE) #must delete first line of C: saving location in txt file
short <- c("Subject", "Comp1", "Comp2", "Distance")
distance <- distance[short]

distance <- na.omit(distance)
distance$Subject <- as.factor(distance$Subject)
distance$Comp1 <- as.factor(distance$Comp1)
distance$Comp2 <- as.factor(distance$Comp2)

posSPAMmood <- c("CALM", "CAREFREE", "CHEERFUL", "COMPOSED", "CONTENT", "EXCITED", "HAPPY", "JOYFUL", "LIVELY", "MANIC", "OPTIMISTIC", "PASSIONATE", "RELAXED", "SATISFIED", "SERENE")
negSPAMmood <- c("ANXIOUS", "BITTER", "BORED",  "DEPRESSED", "DESPERATE", "FED UP", "GLOOMY", "GUILTY", "IRRITABLE", "MISERABLE", "PANICKY", "PESSIMISTIC", "SAD", "STRESSED", "TENSE", "UNEASY")
distance <- distance %>% 
  rename_with(.cols = 1, ~"PtcptID")
distancetest$termpairs <- paste(distancetest$Comp1, distancetest$Comp2, sep= "-")

# REMOVE BORED (outlier term)
distancetest <- distancetest[!grepl("BORED", distancetest$Comp1),]
distancetest <- distancetest[!grepl("BORED", distancetest$Comp2),]
```

# Loops for individual matrices, individual configuration plots, stress values (stress WIP)
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)

rownames(distancetest) <- 1:nrow(distancetest)

# INDIVIDUAL MATRIX CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #save all files in Imp folder for use at individual level
    savefile <- paste0 ("~/ownCloud/PhD/Projects/Data/Myriad1B/indivmatrices/", i, ".csv", sep = "")
    write.csv(SpamMatrixtemp, file = savefile, row.names = TRUE)
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}

# INDIVIDUAL GGSCATTER CONFIGURATION PLOT CREATION LOOP
for(i in levels(test0dem$PtcptID)) {
    #single out an individual participant
    matrixloop1 <- distancetest[distancetest$PtcptID == i,]
    an <- with(matrixloop1[,2:4], sort(unique(c(as.character(Comp1), as.character(Comp2)))))
    SpamMatrixtemp <- array(0, c(length(an), length(an)), list(an, an))
    a <- match(matrixloop1[,2:4]$Comp1, an)
    b <- match(matrixloop1[,2:4]$Comp2, an)
    SpamMatrixtemp[cbind(a,b)] <- SpamMatrixtemp[cbind(b,a)] <- matrixloop1[,2:4]$Distance
    #create their mds plot
    spam_mdstemp <-  mds(delta = SpamMatrixtemp, ndim = 2, type = "ordinal")
  #simple clean readable config plot
  plottemp <- ggplot() + geom_point(data = as.data.frame(spam_mdstemp$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdstemp$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(SpamMatrixtemp)) + 
  labs(title = "MDS representation of Mood") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()

    #save all files in Imp folder for use at individual level
    savefile <- paste0 (i, ".jpg", sep = "")
    ggsave(file=savefile, plot=plottemp, path = "~/ownCloud/PhD/Projects/Data/Myriad1B/indivmatrices/indivconfigplots_NoBored/")
    #print output into console to see status
    print(paste("Dataframe Saved:", i))
    #save outputs to location with unique name match.
}
```

# Average Low and At Risk Group analysis of matrices (averaging matrices, stress, plots, variance diff)
```{r}
library(data.table)
library(dplyr)
library(reshape2)
library(smacof)
library(magrittr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(stats)

#Average matrices for low and at risk groups by SUBFOLDER OF PRESORTED GROUPS.
    ## LOW RISK ##
#pull all the csv files from the subfolder
filenamesLR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatrices/lowrisk_noBored", pattern="*.csv", full.names=TRUE)
ldfLR <- lapply(filenamesLR, read.csv)
resLR <- lapply(ldfLR, as.matrix)
#make each one have their mood terms as row names via lapply.
resLR <- lapply(ldfLR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
#matrix average LOW RISK
AvgMatrixLoRisk <- Reduce(`+`, resLR)/length(resLR)

    ## AT RISK ##
filenamesHR <- list.files(path ="~/Desktop/Projects/Data/Myriad1B/indivmatrices/atrisk_noBored", pattern="*.csv", full.names=TRUE)
ldfHR <- lapply(filenamesHR, read.csv)
resHR <- lapply(ldfHR, as.matrix)
#make each one have their mood terms as row names via lapply.
resHR <- lapply(ldfHR, function(df) {
    #new df without rowname col
    df_out <- df[,-1]
    #set rownames as first col from input df
    rownames(df_out) <- df[[1]]
    df_out
})
AvgMatrixHiRisk <- Reduce(`+`, resHR)/length(resHR)

#Assess group matrices 
#analyze configuration plots, stress values, normalized stress, and k-means clustering of outputs.

#Low Risk
spam_mdsLR <-  mds(delta = AvgMatrixLoRisk, ndim = 2, type = "ratio" )
spam_mdsLR
spam_mdsLR$stress #the stress value = .13 GREAT
dhat_matrix <-  as.matrix(spam_mdsLR$dhat)
d_matrix <-  as.matrix(spam_mdsLR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressLR  <-  nominator/denominator
normalized_stressLR #the normalized stress value = 0.02

#At Risk
spam_mdsHR <-  mds(delta = AvgMatrixHiRisk, ndim = 2, type = "ratio" )
spam_mdsHR
spam_mdsHR$stress #the stress value = .13 GREAT
dhat_matrix <-  as.matrix(spam_mdsHR$dhat)
d_matrix <-  as.matrix(spam_mdsHR$confdist)
denominator <-  sum(dhat_matrix[upper.tri(dhat_matrix)]^2)
p_ij <-  dhat_matrix[upper.tri(dhat_matrix)]
d_ij <-  d_matrix[upper.tri(d_matrix)]
nominator <-  sum((p_ij - d_ij)^2) 
normalized_stressHR  <-  nominator/denominator
normalized_stressHR #the normalized stress value = 0.02

#compare spread of both dimensions.
LRscores <- as.data.frame(spam_mdsLR$conf)
mean(LRscores$D1)
sd(LRscores$D1)
mean(LRscores$D2)
sd(LRscores$D2)

MoodLRscores <- LRscores

HRscores <- as.data.frame(spam_mdsHR$conf)
mean(HRscores$D1)
sd(HRscores$D1)
mean(HRscores$D2)
sd(HRscores$D2)

MoodHRscores <- HRscores

#Group plots by risk status
plot(spam_mdsLR)
plot(spam_mdsHR)
#Low risk
LRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsLR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsLR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixLoRisk)) + 
  labs(title = "MDS representation of Mood: Low Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()
    ggsave(file="LRplotnoBoredRatio.jpg", plot=LRplot, path = "~/Desktop/Projects/Data/Myriad1B/")
#At risk
HRplot <- ggplot() + geom_point(data = as.data.frame(spam_mdsHR$conf), 
                      mapping = aes(x = D1 , y = D2), 
                      alpha = 0.5 , color = "blue", size = 3) + 
  geom_text_repel(data = as.data.frame(spam_mdsHR$conf), 
            mapping = aes(x = D1,y= D2), 
            label = rownames(AvgMatrixHiRisk)) + 
  labs(title = "MDS representation of Mood: At Risk Group") + xlab("Dimension 1") + ylab ("Dimension 2") +
  theme_classic()
  #scale_y_continuous(limits = c(-.5, .4)) #make same as low risk plot Y (X is already same)
    ggsave(file="HRplotnoBoredRatio.jpg", plot=HRplot, path = "~/Desktop/Projects/Data/Myriad1B/")


#t test for mean differences
t.test(LRscores$D1, HRscores$D1)
#t = -2.6686e-16, df = 59.955, p-value = 1
t.test(LRscores$D2, HRscores$D2)
#t = -4.859e-15, df = 49.249, p-value = 1

#ð¹-test of  two variances
var.test(LRscores$D1, HRscores$D1)
var.test(LRscores$D2, HRscores$D2)

LRscores$Group <- "LowRisk"
HRscores$Group <- "AtRisk"

LRHR <- rbind(LRscores, HRscores)
#plot the spread
#boxplot(D2 ~ Group, data = LRHR)
ggplot(LRHR, aes(x=Group, y=D1, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 1 between Groups") +
  theme_classic()
ggplot(LRHR, aes(x=Group, y=D2, fill=Group)) + geom_boxplot() + ggtitle("Spread of Dimension 2 between Groups") +
  theme_classic()

#within-groups variance of dimensions
var.test(LRscores$D1, LRscores$D2)
var.test(HRscores$D2, HRscores$D2)


# STATISTIC FOR ASSESSING BORED OUTLIER
checkingoutliers <- as.data.frame(spam_mdsLR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)
z_scores

checkingoutliers <- as.data.frame(spam_mdsHR$conf)
#get Z-value for the D2 variables.
checkingoutliers$z_scores <- (checkingoutliers$D2-mean(checkingoutliers$D2))/sd(checkingoutliers$D2)
z_scores
```

#Clustering attempts, difference in groups
```{r}
library(stats)
library(dplyr)
library(ggplot2)
library(ggpubr)
mdsLR <- spam_mdsLR$conf
mdsHR <- spam_mdsHR$conf

# Determine number of clusters
wss <- (nrow(mdsLR)-1)*sum(apply(mdsLR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsLR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="Low Risk Group") #strongly 2

wss <- (nrow(mdsHR)-1)*sum(apply(mdsHR,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mdsHR,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main ="At Risk Group") #also strongly 2

# Ward Hierarchical Clustering
d <- dist(mdsLR, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward")
plot(fit) # display dendogram
groups <- cutree(fit, k=4) # cut tree into N clusters
# draw dendogram with red borders around the N clusters
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")

d <- dist(mdsHR, method = "euclidean")
fit <- hclust(d, method="ward")
plot(fit)
groups <- cutree(fit, k=4)
rect.hclust(fit, k=2, border="blue")
rect.hclust(fit, k=4, border="red")
```
